% Parse Rate Transparency — for inclusion in Method section
% Addresses Reviewer 3's concern about unreported parsing success rates

\paragraph{Response Parsing and Missing Data.}
Each LLM response was parsed by extracting the final numeric value ($\in [0,1]$) from the model's free-text output. Table~\ref{tab:parse-rates} reports completion and parse rates for the model survey.

In Phase~1 (prompt screening, Gemini~2.0~Flash), all 14{,}700 expected responses were received and parsed successfully (100\% across 35 conditions). In Phase~2 (model survey), completion varied by model. Gemini and Claude achieved 100\% on all attempted prompts. Locally-hosted Llama models achieved 93\% completion. GPT-4o was the most problematic: only 66.7\% of expected files were present and 65.4\% parseable---primarily due to \emph{contrastive} and \emph{simulation} conditions where the model produced narrative responses without a final numeric estimate. This means GPT-4o's reported correlations (Table~\ref{tab:models}) are based on a non-random subset of items, and may not be directly comparable to other models. Qwen3-32B parsed at only 37.5\%, confirming our decision to exclude it.

Correlations are computed over successfully parsed items only; unparseable responses are treated as missing data, not imputed.

\begin{table}[t]
\centering
\caption{Phase~2 parse rates by model (aggregated across all prompt framings).}\label{tab:parse-rates}
\small
\begin{tabular}{lrrrrr}
\toprule
\textbf{Model} & \textbf{Expected} & \textbf{Present} & \textbf{Parsed} & \textbf{File\%} & \textbf{Parse\%} \\
\midrule
Gemini 2.0 Flash      & 2{,}100 & 2{,}100 & 2{,}100 & 100.0 & 100.0 \\
Claude Opus 4.5       & 1{,}260 & 1{,}260 & 1{,}260 & 100.0 & 100.0 \\
GPT-4o                & 2{,}100 & 1{,}400 &    915  &  66.7 &  65.4 \\
Llama 3.3 70B         & 2{,}100 & 1{,}949 & 1{,}948 &  92.8 &  99.9 \\
Llama 3.1 8B          & 2{,}100 & 1{,}948 & 1{,}918 &  92.8 &  98.5 \\
Llama 4 Scout         & 2{,}100 & 1{,}955 & 1{,}952 &  93.1 &  99.8 \\
Gemma 3 27B           & 2{,}100 & 1{,}502 & 1{,}502 &  71.5 & 100.0 \\
Qwen 3 32B (excluded) & 1{,}260 &    472  &    426  &  37.5 &  90.3 \\
\bottomrule
\end{tabular}
\end{table}
